# Data {#appendixdata .unnumbered}

TODO fill in

## hcandersenr {-}

The **hcandersenr**[@R-hcandersenr] package includes the text of the 157 known fairy tales by the Danish author H.C. Andersen. 
The text comes with 5 different languages with 

- 156 in English,
- 154 in Spanish,
- 150 in German,
- 138 in Danish and
- 58 in French

The package comes with a dataset for each language with the naming convention `hcandersen_**`,
where `**` is a country code.
Each dataset comes as a data.frame with two columns; `text` and `book` where the `book` variable has the text divided into strings of up to 80 characters.

The package also comes with a dataset called `EK` which includes information about the publication date, language of origin and names in the different languages.

## scotus {-}

The **scotus**[@R-scotus] package contains a sample of the Supreme Court of the United States' opinions.
The `scotus_sample` data.frame includes 1 opinion per row along with the year, case name, docket number, and a unique ID number.

The text has had minimal preprocessing done on them and will include the header information in the text field.
Example of the beginning of an opinion is shown below

```{r, echo=FALSE}
library(scotus)
cat(paste(stringr::str_wrap(head(strsplit(scotus_sample$text[42], "\n")[[1]])), collapse = "\n"))
```

## GitHub issue {-}

This [dataset](https://github.com/explosion/projects/tree/master/textcat-docs-issues) includes 1161 Github issue title and an indicator of whether the issue was about documentation or not.
The dataset is split into a training data set and evaluation data set.

TODO find out how to store this data

```{r, eval=FALSE}
library(jsonlite)
library(readr)
library(tidyverse)

json_to_df <- function(x) {
  json <- parse_json(x)
  
  tibble(text = json$text,
         label = json$label,
         answer = json$answer)
}

github_issues_training <- read_lines("https://raw.githubusercontent.com/explosion/projects/master/textcat-docs-issues/docs_issues_training.jsonl") %>%
  map_dfr(json_to_df)

github_issues_eval <- read_lines("https://raw.githubusercontent.com/explosion/projects/master/textcat-docs-issues/docs_issues_eval.jsonl") %>%
  map_dfr(json_to_df)
```

